Recurrent event analysis in the presence of real-time high frequency data via random subsampling

### 2020-01-27 write-up/fda-recurrent.tex
tex updating!
Adding changes from SPH paper
Edited simulations and paper.

\begin{assumption}[Event process assumptions]
\label{assumption:events}\normalfont
We assume the following holds:
\begin{enumerate}[label=(E.\arabic*)]
\item\label{E1} The subsampling rate is both lower and upper bounded for all at-risk times; that is,~$0 < L < \pi_i (t) < U < \infty$ for all~$i=1,2,\ldots$ and~$t \in [0, \tau]$ such that $R_i (t) = 1$.
\item\label{E1} The subsampling rate is both lower and upper bounded for all ``at risk times''; that is,~$0 < L < \pi_i (t) < U < \infty$ for all~$i=1,2,\ldots$ and~$t \in [0,\tau]$.
\item\label{E2} There exists a nonnegative definite matrix~$\Xi (\theta)$ such that
  \[
    n^{-1} \Xi_n (\theta) = n^{-1} \sum_{i=1}^n \int_0^\tau w_i (t;
    \theta) (1 - w_i (t; \theta)) \left[ \frac{h_i^{(1)}(t;
        \theta)}{h_i (t; \theta)} \right] \times  \left [
      \frac{h_i^{(1)} (t;\theta)}{h_i(t; \theta)} \right]^\top R_i (t)
    dt \overset{P}{\to} \Xi (\theta).
  \]
\item\label{E3} There exists~$M$ such that $|W_{i,j,t}| < M$ for all~$(i,j,t)$.
\item\label{E4} For all~$j,k$
\[
n^{-1} \sum_{i=1}^n \int_0^\tau \left | \frac{d^2}{d\theta_j
    d\theta_k} h_i (t;\theta_0) \right|^2 R_i (t) dt \overset{P}{\to}
C < \infty
\]
as $n \to \infty$.
\end{enumerate}
\end{assumption}

@@

We next assess the proposed methodology via a simulation study. Here, we assume each individual is observed over five days each defined on the unit interval~$[0,1]$ with $1000$ equally spaced observation times per day. We define~$X(t)$ at the grid of observerations as a mean-zero Gaussian process with covariance
\[
  \Sigma (t, t^\prime) =
  \frac{\sigma^2}{\Gamma (v) 2^{\nu-1}}
  \left( \frac{\sqrt{2 \nu} |t-s|}{\rho_t} \right)^{\nu}
  K_v \left( \frac{\sqrt{2 \nu} |t-s|}{\rho_t} \right)
\]
where~$K_v$ is the modified Bessel function of the second kind.  We set~$\nu = 1/2$, $\sigma^2 = 1$, and~$\rho = 0.5$ as well as set~$K_b = K_x = 35$.  For simplicity, we assume~$\Sigma$ is known in computation of the the eigendecomposisions. Given~$\{ X(t) \}_{0 \leq t \leq 1}$ for a given user-day, we generate event times according a chosen hazard function~$h (t; theta)$.  To mimic our real data, we set
\[
h(t; \theta) = \exp \left( \theta_0 + \int_{0}^{\Delta} X(t-s) \beta(s) ds \right).
\]
We set~$\Delta = 1/(12*2)$ to mimic a $30$-minute window for a $12$-hour day.  We set~$\theta_0 = \log(5/1000)$ to set a baseline risk of approximately 5 events per day. We consider two choices of~$\beta(s)$: (1) $\beta_0 + \exp(- \beta_1 s) $, and (2) $\beta_1  * \sin \left( 2 \pi \frac{s}{\Delta} - \pi/2 \right)$

@@

\begin{table}[!th]
\begin{tabular}{c | c c c c c}
Sampling rate & 0.5 & 1 & 5 & 25 & 50 \\
Variance
Squared Bias
MISE
\end{tabular}
\caption{XXX}
\label{tab:mise}
\end{table}

@@

In each (A)-(C) we parametrize the curves and check the type I error
for various sizes of the

@@

Theorem~XX shows that if we keep~$K_x = K_b = 35$ and therefore have reconstruction error, we can expect a downward bias in parameter estimates. Moreover, if~$\Delta < \Delta^\star$ then we can expect a similar bias in the parameter estimates.  For the former, however, we can simply let~$K_x = K_b$ grow as a function of data size; for the latter, we can see if~$\hat \beta(s)$ is statistically significant for~$s$ near~$\Delta$. If so, we can set~$\Delta$ larger to ensure the condition holds.

@@

@@

During a XX month period in 2019, 100 psychiatric inpatients admitted
for suicidal risk to Franciscan Children's Hospital

hospitals with histologically verified liver cirrhosis were recu
those presenting to a psychiatric hospital with suicide ideation
and/or a recent suicide attempt) during a high risk time period (i.e.,
post-hospitalization)

Currently, only 51 patients for whom the initial biopsy could be
reevaluated using more restrictive criteria were available for
analysis.
The goal of the study was to ask whether sensor information can used
to identify distress early prior in the process of escalation by
monitoring physiological correlates of distress -- in particular,
electrodermal activity (EDA), a measure of skin conductance, and
``gACC'', a coordinate-free feature built from accelerometer data
that measures physical movement.
The scientific hypothesis to be tested is whether a spike in
EDA and/or gACC

@@

@@

To prove~\eqref{eq:andersen3},
\[
\left | n^{-1}  \right | \leq \frac{1}{n} \sum_{i=1}^n \int_0^\tau H_{in} (t) dN_i^
\]
\end{proof}

>>>>

\begin{assumption}[Event process assumptions]
\label{assumption:events}\normalfont
We assume the following holds:
\begin{enumerate}[label=(E.\arabic*)]
\item\label{E1} The subsampling rate is both lower and upper bounded for all at-risk times; that is,~$0 < L < \pi_i (t) < U < \infty$ for all~$i=1,2,\ldots$ and~$t \in [0,\tau]$ such that $R_i (t) = 1$
\item\label{E2} There exists a nonnegative definite matrix~$\Xi (\theta)$ such that
  \[
    n^{-1} \Xi_n (\theta) = n^{-1} \sum_{i=1}^n \int_0^\tau w_i (t;
    \theta) (1 - w_i (t; \theta)) \left[ \frac{h_i^{(1)}(t;
        \theta)}{h_i (t; \theta)} \right] \times  \left [
      \frac{h_i^{(1)} (t;\theta)}{h_i(t; \theta)} \right]^\top R_i (t)
    dt \overset{P}{\to} \Xi (\theta).
  \]
\item\label{E3} There exists~$M$ such that $|W_{i,j,t}| < M$ for all~$(i,j,t)$.
\item\label{E4} For all~$j,k$
\[
n^{-1} \sum_{i=1}^n \int_0^\tau \left | \frac{d^2}{d\theta_j
    d\theta_k} h_i (t;\theta_0) \right|^2 R_i (t) dt \overset{P}{\to}
C < \infty
\]
as $n \to \infty$.
\end{enumerate}
\end{assumption}

@@

We next assess the proposed methodology via a simulation study. Here, we assume each individual is observed over five days where each day is defined on the unit interval~$[0,1]$ with $1000$ equally spaced observation times per day. We define~$X(t)$ at the grid of observations as a mean-zero Gaussian process with covariance
\[
  \Sigma (t, t^\prime) =
  \frac{\sigma^2}{\Gamma (v) 2^{\nu-1}}
  \left( \frac{\sqrt{2 \nu} |t-s|}{\rho_t} \right)^{\nu}
  K_v \left( \frac{\sqrt{2 \nu} |t-s|}{\rho_t} \right)
\]
where~$K_v$ is the modified Bessel function of the second kind.  We set~$\nu = 1/2$, $\sigma^2 = 1$, and~$\rho = 0.3$ as well as set~$K_b = K_x = 35$.  For simplicity, we assume~$\Sigma$ is known in computation of the eigendecompositions. Given~$\{ X(t) \}_{0 \leq t \leq 1}$ for a given user-day, we generate event times according a chosen hazard function~$h (t; \theta)$.  To mimic our real data, we set
\[
h(t; \theta) = \exp \left( \theta_0 + \int_{0}^{\Delta} X(t-s) \beta(s) ds \right).
\]
We set~$\Delta = 1/(12*2)$ to mimic a $30$-minute window for a $12$-hour day.  We set~$\theta_0 = \log(5/1000)$ to set a baseline risk of approximately 5 events per day. We consider two choices of~$\beta(s)$: (1) $\beta_0 + \exp(- \beta_1 s) $, and (2) $\beta_1  * \sin \left( 2 \pi \frac{s}{\Delta} - \pi/2 \right)$

@@

\begin{table}[!th]
\begin{tabular}{c | c c c c c}
Sampling rate & 0.5 & 1 & 5 & 25 & 50 \\ \hline
Variance & (10,12) & - & - & - & - \\
Squared Bias & - & - & - & - & -  \\
MISE & - & - & - & - & -  \\ \hline
Type I error & - & - & - & - & -  \\ \hline
\end{tabular}
\caption{Mean-integrated squared error and type I error as a function of sampling rate for $\beta(s)$ given by (1) and (2) respectively.}
\label{tab:mise}
\end{table}

@@

We simulate the system $1000$ times under $\beta \equiv 0$ and check the type I error.

@@

Theorem~XX shows that if we keep~$K_x = K_b = 35$ and therefore have reconstruction error, we can expect a downward bias in parameter estimates. Moreover, if~$\Delta < \Delta^\star$ then we can expect a similar bias in the parameter estimates.  For the former, however, we can simply let~$K_x = K_b$ grow as a function of data size; for the latter, we can see if~$\hat \beta(s)$ is statistically significant for~$s$ near~$\Delta$. If so, we can set~$\Delta$ larger to ensure the condition holds.  In each simulation, we use $\Delta \in (18,21,24,27,30)$.

\begin{table}[!th]
\begin{tabular}{c c c c c c}
& \multicolumn{5}{c}{Sampling rate} \\ \cline{2-6}
$\Delta=$ & 0.5 & 1 & 5 & 25 & 50 \\ \hline
18  & - & - & - & - & -  \\
21 & - & - & - & - & -  \\
24 & - & - & - & - & -  \\
27 & - & - & - & - & -  \\
30 & - & - & - & - & -  \\ \hline
\end{tabular}
\caption{XXX}
\label{tab:mise}
\end{table}

@@

\subsection{Online estimation}

@@

During a XX month period in 2019, 91 psychiatric inpatients admitted for suicidal risk to Franciscan Children's Hospital  were enrolled in
 with histologically verified liver cirrhosis were recu
those presenting to a psychiatric hospital with suicide ideation
and/or a recent suicide attempt) during a high risk time period (i.e.,
post-hospitalization)

Currently, only 51 patients for whom the initial biopsy could be
reevaluated using more restrictive criteria were available for
analysis.
One of the goals of the study was to assess the association between sensor-based covariate information can used to identify distress early prior in the process of escalation by monitoring physiological correlates of distress -- in particular, electrodermal activity (EDA), a measure of skin conductance, and ``gACC'', a coordinate-free feature built from accelerometer data that measures physical movement. The scientific hypothesis to be tested is whether a spike in EDA and/or gACC

@@

{\bf comparison over $\Delta$ and using imputation}

@@

To prove~\eqref{eq:andersen3},
\[
\left | n^{-1}  \right | \leq \frac{1}{n} \sum_{i=1}^n \int_0^\tau H_{in} (t) dN_i
\]
\end{proof}
